The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-08-11 Independence_CH/related.tex
Minor edits

It should be emphasized, however, that the Flypitch project was
carried out assuming a rather strong meta-theory. 
Carneiro \cite{carneiro-ms-thesis} reports that Werner's results in
\cite{10.5555/645869.668660} can be adapted to show that the base
logic  of Lean (restricted to $n$
type universes) proves the consistency of $\ZFC$ plus $n$
inaccessibles. Han and van Doorn did use universes in their
implementation; for instance, ordinals are “defined as equivalence
classes of (well-ordered) types, [\dots] one
universe level higher than the types used to construct them”
\cite{han_et_al:LIPIcs:2019:11074}. It is not clear to us if they are
able to avoid such strength (at least $\Con(\ZFC)$ is readily
available in order to get either unprovability result).
% by Gödel's Second).

On a lesser note, in order to prove $\AC$ in the generic extension,
Flypitch requires choice in the metatheory
\cite[p.~11]{han_et_al:LIPIcs:2019:11074}. Our formalization works
entirely in $\ZF$, and only import the Isabelle theory postulating
$\AC$ for demostration purposes.

>>>>

It should be emphasized, however, that the Flypitch project was
carried out assuming a rather strong meta-theory. 
Carneiro \cite{carneiro-ms-thesis} reports that Werner's results in
\cite{10.5555/645869.668660} can be adapted to show that the base
logic  of Lean (restricted to $n$
type universes) proves the consistency of $\ZFC$ plus $n$
inaccessibles. Han and van Doorn did use universes in their
implementation; for instance, ordinals are “defined as equivalence
classes of (well-ordered) types, [\dots] one
universe level higher than the types used to construct them”
\cite{han_et_al:LIPIcs:2019:11074}. It is not clear to us if they are
able to avoid such strength (at least $\Con(\ZFC)$ is provable in
their context, a requisite to get either unprovability result).
% by Gödel's Second).

On a lesser note, in order to prove $\AC$ in the generic extension,
Flypitch requires choice in the metatheory
\cite[p.~11]{han_et_al:LIPIcs:2019:11074}. Our formalization works
entirely in $\ZF$, and only imports the Isabelle theory postulating
$\AC$ for demostration purposes.
